---
title: "My R Shortcut library"
author: "Nico Steckhan"
date: "15 8 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

Codeing can be a time consuming task. Therfore I created a library where I collect code snippets to have shortcuts right to my fingertips.



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Generate demo data
```{r demodata, echo=FALSE}
data <- data.frame(id=1:100,group=sample(c("verum","control"),100,replace = T),time=rep(1,100),output=runif(100))
data2 <- data.frame(id=1:100,group=sample(c("verum","control"),100,replace = T),time=rep(2,100),output=runif(100))

data <- rbind(data,data2)

```

## repeat to fill column with ID or group ####
```{r pressure, echo=FALSE}
id <- na.omit(data$group)
data$group <- rep(id, each=2)
```



## fill group column (if Visiten have NAs, eg just first group in visite 1 is mentioned) #### 
df <- data.frame(na.omit(data$Gruppe))
data$Gruppe <- df[rep(seq_len(nrow(df)), each=2),] # 2 Visiten

##### remove id if any timepoint is NA (for ezANOVA) ###########
setwho2
setwho2 <- na.omit(setwho2)
setwho2 <- subset(setwho2, Visite != "2") ## keine Visite 2
setwho2 <- subset(setwho2, Visite != "3") ## keine Visite 3
##### keep id if all roundwards are available ######
vis1 <- c(0,1) # AND keep only complete Visiten 1, 2
x1 <- do.call(rbind, lapply(split(data, list(data$VPNR)), 
                            function(.var){ 
                              if (all(.var$Visite == vis1)) return(.var) # or %in%
                              else return(NULL) 
                            })) 



## comma into dot ####
data[,"BMI"] <- gsub(",", ".", data[ , "BMI"])  # einzeln pro spalte
data <- data.frame(lapply(data, function(x) gsub(",", ".", x, fixed = TRUE))) #für alle Spalten im dataframe

## from wide to long format (for ggplot2) ####
# http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/

library(reshape2)

fastdia2 <- na.omit(fastdia2)
fastdia2$sub <- rownames(fastdia)
data_long <- melt(fastdia2,
                  # ID variables - all the variables to keep but not split apart on
                  id.vars=c("id"),
                  # The source columns
                  measure.vars=c("Homa0", "Homa1"),
                  # Name of the destination column that will identify the original
                  # column that the measurement came from
                  variable.name="time",
                  value.name="measurement"
)


datal <- reshape(data[1:40], idvar=c("PID"), varying=2:40, direction="long", sep="_")
boxplot(KG ~ time , data=datal)

library(reshape2)
melt(data, id.vars=c("PID"))



library(tidyr)
data1 <- data[1:4]
gather(data1, time, KG, 2:4, factor_key=T)
data1l <- gather(data1, time, KG, 2:4, factor_key=T)
boxplot(KG ~ time, data = data1l)

data[,1:10] %>%
  gather(key, value, -PID) %>%
  extract(col = key,into =  c("variable", "time"), regex = "([A-z]+)_([A-z0-9]+)") %>%
  spread(variable, value)





## merge data ''
data3 <- merge(data, data_long, by=c("VPNR","Visite"))
## merge by rownames 
merge(entspanntab,entsptab,by="row.names",all.y=TRUE)

### multiple imputation with amelia --------------------------------------
library(Amelia)
ncpus = 4 #for multicore support
m = ncpus * 10      # 40 sets, see Bagging
summary(x1)
a.out <- amelia(x=x1,m=m, idvars = "VPNR", ts="Visite", cs="Gruppe",
                ords = c("SF01_01","SF02","SF03_03","SF05_01"),  parallel = 'multicore', ncpus = ncpus)
a.out$imputations[[5]]




# restore order 
x.new$Pat_Nr <- as.numeric(as.character(x.new$Pat_Nr))
x.new <- x.new[order(x.new[,1]),]

# as.numeric(as.character(x)) shortcut
anac <- function (x) {x <- as.numeric(as.character(x))}

# split name and edit
rownames(entsptab) <- gsub("^.*?ENTSP_","", rownames(entsptab))
# or split like this
save=NULL
for (i in 1:length(rownames(entsptab))) {
        save <- append(save, print(unlist(strsplit(rownames(entsptab)[i], split='_', fixed=TRUE))[4]))
}
save
rownames(entsptab) <- save 
#####
### Build spliting variable from 
library(Hmisc)
cut2(rm1.complete$expr, g=2)


# Check if Variable has NA across Visits ----------------------------------

table(is.na(data$Gew),data$Visite)

as.data.frame(split(ifelse(is.na(data$Gew)==TRUE,NA,1), data$VPNR),optional = T)
check <- t(as.data.frame(split(ifelse(is.na(data$Gew)==TRUE,NA,1), data$VPNR),optional = T))
md.pattern(check)
# scan thru columns to find NA patterns -----------------------------------
ba <- colnames(data)
for (i in ba) {
  chek <- t(as.data.frame(split(ifelse(is.na(data[,i])==TRUE,NA,1), data$VPNR),optional = T))
  print(i)
  #print(chek)
  print(md.pattern(chek))
}


# Loop thru table an build Word-format tables
# Ala MIFT.R





# good 2 know -------------------------------------------------------------

# aggregate multiple functions
aggregate(iris$Sepal.Length, by=list(iris$Species), function(x){c(mean=mean(x),sdev= sd(x)) })


# Wilcoxn effect size (output r)-----------------------------------------------------

wilcox_effectsize <- function (x,n){
  require(coin)
  z = statistic(wilcox_test(x, distribution='exact'))
  return(abs(z/sqrt(n)))
}
 

# add column inbetween ----------------------------------------------------
as.data.frame(append(df1, df2, after = 3))



# Reorder factor levels ---------------------------------------------------
data$Gruppe <- factor(data$Gruppe, levels = levels(data$Gruppe)[c(2,1)])


# list all means etc ----------------------------------------------------
data4_describe = matrix(unlist(apply(data4,2,describe)),ncol=13 , byrow = T)
write.clipboard(cbind(names(data4),data4_describe))


# extract every nth element of a vector -----------------------------------
a <- 1:120
b <- a[seq(1, length(a), 6)]


# bind function to variables ----------------------------------------------
makeActiveBinding('x', function () dim(abf), globalenv())


# remove whitespace in characters
names(data) <- gsub(" ", "", names(data), fixed = TRUE)

# Extract se cond subelement of every element in a list while ignoring NA's in sapply
sapply(mylist,function(x) x[2])



# Multiplot of different kinds --------------------------------------------

library(RGraphics) # support of the "R graphics" book, on CRAN
library(gridExtra) 
g1 <- tableGrob(head(iris))
string <- "
This famous (Fisher's or Anderson's) iris data set gives the
measurements in centimeters of the variables sepal length and width
and petal length and width, respectively, for 50 flowers from each of
3 species of iris. The species are Iris setosa, versicolor, and
virginica.
"
g2 <- splitTextGrob(string)
#"Sepal.Length" "Sepal.Width"  "Petal.Length" "Petal.Width"  "Species"
g3 <- qplot(Sepal.Length,  Petal.Length, data=iris, colour=Species)
grid.arrange(g1,g2,g3, ncol=1, top="The iris data")



# win to unix -------------------------------------------------------------

# in terimnal type tr -d '\15\32' < winfile.txt > unixfile.txt





# loop thru data for repeated 2-way anovas --------------------------------

myfunc <- function(the_data, the_dv) {
  eval(
    substitute(
      ezANOVA(data = the_data, 
              wid = VPNR, 
              dv = the_dv, 
              between = Gruppe,
              within=Visite,
              type=3,
              detailed = T), 
      list(the_dv = the_dv)))
}


for (i in 74:98) {
  print(colnames(x1)[i])
  print(myfunc(x1,colnames(x1)[i]))
}


# select columns via GUI
vars <- select.list(names(data),multiple=TRUE,
                    title='select your variable names',
                    graphics=TRUE)

# update R interactivley
install.packages("devtools")
devtools::install_github("AndreaCirilloAC/updateR")
updateR(admin_password = "PASSWORD") # Where "PASSWORD" 


# 95% confidence interval into standard deviation
int2sd <- function(a,b,n) {
  ci <- c(a,b)
  x_bar <- mean(ci)
  if (n <= 30) {
    warning("attention: probably non-gaussian")
  }
  return(((sqrt(n))/1.96)*(x_bar-ci[1]))
} 


# SD change
sd_change <- function(sd1,sd2,cor=0.7) {
  return(sqrt(sd1^2+sd2^2 - (2*cor*sd1*sd2)))
}

# SD change from two 95%-intervals
sd_change(sd1=int2sd(-1,2,40),sd2=int2sd(-1.5,1.8,45))


# cut into factor using quantiles -----------------------------------------
alt <- rnorm(n = 100, 1)
cut(alt, breaks = c(quantile(alt, probs =seq(0,1,by=.25))),labels=c("0-25%","25-50%","50-75%","75-100%"), include.lowest = T)

# unlist list
idsn = unlist(lapply(idsn, `[[`, 1))

# copy from console to build vector commaseperated
cat(paste(shQuote(colnames(final), type="cmd"), collapse=", "))

# get datetime format correct (?strptime)
as.POSIXct("Mo 16 Jul 2018 22:58:49", format="%a %d %b %Y %H:%M:%S")
strptime("Mo 16 Jul 2018 22:58:49", format="%a %d %b %Y %H:%M:%S")

# compare datetime
format(as.POSIXct(seesel$V1,format="%H:%M:%S")) > format(as.POSIXct("8:00:00", format="%H:%M:%S"))

# write psych output to clipboard or email
data4_describe = matrix(unlist(apply(data3[,vars <- select.list(names(data3)[1:10],multiple=TRUE,
                                                                title='select your variable names',
                                                                graphics=TRUE)],2,describe)),ncol=13 , byrow = T)
write.clipboard(cbind(names(data3)[names(data3) %in% vars],data4_describe))

# use with (feste breite font: )
knitr::kable(see)

# as html
output <- 
  matrix(paste("Content", LETTERS[1:16]), 
         ncol=4, byrow = TRUE)

# https://cran.r-project.org/web/packages/htmlTable/vignettes/tables.html
library(htmlTable)
htmlTable(output)



